Malacology: A Programmable Storage System

### 2016-10-19 paper/paper.tex
paper: generalize cls table, re-align text blocks, fix service table (#99)
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

%\documentclass[10pt,twocolumn]{article}
\documentclass[preprint]{sigplanconf-eurosys}

%\usepackage{fullpage}
%\usepackage{times}
\usepackage{multirow}
\usepackage{arydshln}
%\usepackage{amssymb,amsmath}
%\usepackage{ifxetex,ifluatex}
\usepackage{amsmath}
%\usepackage{mathptmx}
\usepackage{graphicx}
\usepackage{listings}
%\usepackage{amssymb}
\usepackage{pifont}
\usepackage{color}
\newcommand{\cmark}{\ding{51}}%
\usepackage[font=footnotesize,labelfont=bf]{caption}

\newcommand{\paa}[1]{{\textcolor{red}{[[#1 -- paa]]}}}
%\newcommand{\xmark}{\ding{55}}%
%
%
%
%% use upquote if available, for straight quotes in verbatim environments
%\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
%
%\usepackage[font=footnotesize,labelfont=bf]{caption}
\usepackage{pstricks}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
%% Scale images if necessary, so that they will not overflow the page
%% margins by default, and it is still possible to overwrite the defaults
%% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,keepaspectratio}

\usepackage[unicode=true]{hyperref}
\hypersetup{breaklinks=true,
            bookmarks=true,
            pdfauthor={},
            pdftitle={Malacology: A Programmable Storage System},
            colorlinks=true,
            citecolor=blue,
            urlcolor=blue,
            linkcolor=black,
            pdfborder={0 0 0}}
\urlstyle{same}  % don't use monospace font for urls
%%\setlength{\parindent}{0pt}
%%\setlength{\parskip}{6pt plus 2pt minus 1pt}
%%\setlength{\emergencystretch}{3em}  % prevent overfull lines
%\setcounter{secnumdepth}{5}

@@

%These three trends put evolutionary pressure on storage systems and raise the
%question of whether 
The evolutionary pressure placed on storage systems by these trends raises the
question of whether
there are principles that storage systems designers can follow
to evolve storage systems efficiently, without jeopardizing years of
code-hardening and performance optimization efforts.
% that are important for
%users to continue to entrust their data to the storage system.
In this paper we investigate an approach that focuses on 
%generalizing 
%and
%composing 
identifying and exposing existing
storage system resources, services, and abstractions that in a generalized form
can be used to \emph{program} new services. By composing higher-level services
over existing abstraction one can reuse
subsystems and leverage their optimizations, established correctness,
robustness, and efficiency.

@@

{\it \textbf{We implement two distributed services}} using 
%programmable storage
%to demonstrate its feasability:
Malacology to demonstrate the feasibility of the programmable storage approach:

@@

The remainder of this paper is structured as follows. First, we describe and
motivate the need for programmable storage by describing current practices in
the open source community. Next we describe Malacology by presenting the
subsystems within the underlying storage system that we re-purpose, and briefly
describe how those system are used within Malacology (\S\ref{sec:malacology}).
Then we describe the services that we have constructed within the Malacology
framework (\S\ref{sec:services}), and evaluate our ideas within our prototype
implementation (\S\ref{evaluation}).  Finally we discuss related work and
conclude.

@@

% does this belong?
%Ceph~\cite{weil:osdi2006-ceph,ceph}, the increasingly
%popular, production-level open-source distributed storage system, was
%originally built to be an HPC file system. But demands for storage systems that
%could slide in behind cloud infrastructures, like OpenStack, shifted the focus
%to Ceph's internal object store as a backend for large cloud block devices,
%S3-like objects, and file system workloads.

To demonstrate this trend in storage systems we examine the state of
programmability in Ceph. Something of a storage swiss army knife, Ceph simultaneously supports file,
block, and object interfaces on a single
cluster~\cite{ceph_contributors_ceph_2010}. Ceph's Reliable Autonomous
Distributed Object Storage (RADOS) system is a cluster of storage devices
(OSDs) that provide Ceph with data durability and integrity using replication,
erasure-coding, and scrubbing~\cite{weil_rados_2007}. Ceph already 
%has some
provides some degree of
programmability; the OSDs support domain-specific object interfaces that are
implemented by composing existing low-level storage interfaces that execute
atomically. These interfaces are
written in C++ and are statically loaded into the system.

% By introducing programmability concepts into Ceph, we can build new services
% by carefully exposing internal storage services to applications.

@@

\begin{table}[ht]
\centering
  \begin{tabular}{l|l|l}
    Category & Specialization& \# \\ \hline
    \multirow{2}{*}{Locking} & Shared & \multirow{2}{*}{6} \\
                             & Exclusive & \\ \hdashline
    \multirow{3}{*}{Logging} & Replica & 3 \\
                             & State & 4 \\
                             & Timestamped & 4 \\ \hdashline
    \multirow{4}{*}{Metadata Managment} 
                             & RADOS Block Device  & 37 \\
                             & RADOS Gatway & 27 \\
                             & User & 5 \\
                             & Version & 5 \\ \hdashline
    Garbage Collection       & Reference Counting & 4 \\
\end{tabular}
\caption{A variety of RADOS object storage classes exist to expose interfaces
    to applications. \# is the number of methods that use these categories.
}
\label{table:objclasses}
\end{table}

@@

%\section{Challenges}
%\label{sec:challenges}
%
%The challenges of introducing programmability into existing storage systems:
%(1) Programmability while keeping storage system available, (2) Finding and
%factoring out hard-wired policies and optimizations, (3) Bugs that were hidden
%because mechanisms were only exercised according to hard-wired policies and
%not in their full generality they were thought of -- programmability into a
%storage system turned out to be a great debugging tool, (4) version control
%and propagation of updates throughout the system without interrupting
%production workloads, (5) ...

@@

\begin{table*}
\centering
\begin{tabular}{  l | l | l | l }
\multicolumn{4}{c}{\Large \textbf{Common Internal Abstractions}} \\
\multicolumn{4}{c}{} \\
\textbf{Abstraction}                    &
\textbf{Example in Production Systems}  &
\textbf{Example in Ceph}                &
\textbf{Provided Primitives}            \\ \hline
{\it Service Metadata}
  & Zookeeper, Chubby coordination~\cite{hunt_zookeeper_2010,website:chubby}
  & cluster state management~\cite{website:ceph-mon}
  & consensus \& consistency
  \\
{\it Shared Resource}
  & MPI collective IO, burst
  & POSIX metadata protocols~\cite{CITEME}
  & serialization \& batching
  \\
{\it Object \& Data}
  & Swift in situ storage/compute~\cite{website:zerocloud}
  & object interface classes~\cite{website:cls-lua}
  & transaction \& atomicity
  \\
Durability
  & S3/Swift interfaces (RESTful API)
  & object store library~\cite{weil_rados_2007}
  & interface to persistence
  \\
Load Balancing
  & VMWare's VM migration~\cite{vmware-drs,gulati:hotcloud2011-cloud-resource-management} 
  & migration in metadata cluster~\cite{weil:sc2004-dyn-metadata}
  & migration \& sampling
  \\
\end{tabular}
\caption{The same ``internal abstractions" are common across large-scale
systems because they provide primitives that solve general distributed systems
problems.  Here we list examples of what these internal abstractions are used
for in ``production systems" and in Ceph.  Malacology provides these internal
abstractions as interfaces that higher level applications can use; {\it
italicized abstractions} are interfaces contributed in this paper.}
\label{table:examples}
\end{table*}

@@

%Thus, system
%services realized through extensibility should be versioned together as a
%group, exposing the version to allow introspection and domain-specific
%treatment of version mismatch scenarios.

@@

{\bf File types.} The metadata service manages a POSIX file system hierarchy
where files and directories are represented as inode data structures and
expose a POSIX file interface. Application that manage large amounts of
metadata (e.g. users or database snapshots) often require a naming service. We
allow new inode types to be defined such that applications can create
domain-specific interfaces to inodes that may modify locking and capability
policies. We will show this is used in Section~\ref{sec:seq} when we discuss a
distributed shared-log built on Malacology.

{\bf Load balancing.} Mantle~\cite{sevilla:sc15-mantle}, a programmable
metadata load balancer, has been implemented on Malacology so that its
infrastructure can enjoy all the properties of existing internal abstractions.
The load balancing logic is intact but the infrastructure has been improved to
safely load, manage, and persist its load balancing policies.  Composing Mantle
using the service metadata and durability interfaces, Mantle can safely
version balancer policies, save balancer policies in the backend object store
and centralize warnings/errors. When combined with file types, load balancing
policies can express a variety of rules such as multi-tentant and
multi-application balancing.

@@

\label{services}
\begin{table}
\centering
\begin{tabular}{  l | l | l }
\textbf{Service}              &
\textbf{Used to...}           &
\textbf{Malacology Interface} \\ \hline
Mantle & load balance metadata     & Load Balancing  \\
       & agree on current balancer & Versioning      \\ 
       & control client write lock & Shared Resource \\
       & persist policy logic      & Durability      \\ \hline
ZLog   & implement CORFU           & Logical Device  \\ 
       & install ZLog interfaces   & Versioning      \\
       & persist log entries       & Durability      \\ \hline
ZLog   & load balance sequencers   & Load Balancing  \\
Seqr.  & guard shared resource     & Shared Resource \\ 
\end{tabular}
\caption{Mantle and ZLog are built using Malacology interfaces so they inherit the
robustness of the subsystems of the underlying stack.}
\label{table:implementation}
\end{table}

@@

\section{Evaluation}
\label{sec:evaluation}

@@

%\subsubsection{Experiment: Object Class
%Programmability}\label{experiment-object-class-programmability}
%
%Add quote from CORFU paper that talks about what they had to hack
%
%\subsubsection{Experiment: Multi-Client
%Burstiness}\label{experiment-multi-client-burstiness}
%
%\begin{figure}[htbp]
%\centering
%\includegraphics{figures/caps-delay-thruput.png}
%\caption{Forcing the client to drop their capabilities later (delay)
%improves throughput}
%\end{figure}

>>>>

\documentclass[preprint]{sigplanconf-eurosys}
\usepackage{multirow}
\usepackage{arydshln}
\usepackage{amsmath}
\usepackage{graphicx}
\usepackage{listings}
\usepackage{pifont}
\usepackage{color}
\newcommand{\cmark}{\ding{51}}%
\usepackage[font=footnotesize,labelfont=bf]{caption}

\newcommand{\paa}[1]{{\textcolor{red}{[[#1 -- paa]]}}}
\usepackage{pstricks}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
\setkeys{Gin}{width=\maxwidth,keepaspectratio}

\usepackage[unicode=true]{hyperref}
\hypersetup{breaklinks=true,
            bookmarks=true,
            pdfauthor={},
            pdftitle={Malacology: A Programmable Storage System},
            colorlinks=true,
            citecolor=blue,
            urlcolor=blue,
            linkcolor=black,
            pdfborder={0 0 0}}
\urlstyle{same}  % don't use monospace font for urls

@@

The evolutionary pressure placed on storage systems by these trends raises the
question of whether there are principles that storage systems designers can
follow to evolve storage systems efficiently, without jeopardizing years of
code-hardening and performance optimization efforts.  In this paper we
investigate an approach that focuses on identifying and exposing existing
storage system resources, services, and abstractions that in a generalized form
can be used to \emph{program} new services. By composing higher-level services
over existing abstraction one can reuse subsystems and leverage their
optimizations, established correctness, robustness, and efficiency.

@@

{\it \textbf{We implement two distributed services}} using Malacology to
demonstrate the feasibility of the programmable storage approach:

@@

The remainder of this paper is structured as follows. First, we describe and
motivate the need for programmable storage by describing current practices in
the open source community. Next we describe Malacology by presenting the
subsystems within the underlying storage system that we re-purpose, and briefly
describe how those system are used within Malacology (\S\ref{sec:malacology}).
Then we describe the services that we have constructed within the Malacology
framework (\S\ref{sec:services}), and evaluate our ideas within our prototype
implementation (\S\ref{sec:evaluation}).  Finally we discuss related work and
conclude.

@@

To demonstrate this trend in storage systems we examine the state of
programmability in Ceph. Something of a storage swiss army knife, Ceph
simultaneously supports file, block, and object interfaces on a single
cluster~\cite{ceph_contributors_ceph_2010}. Ceph's Reliable Autonomous
Distributed Object Storage (RADOS) system is a cluster of storage devices
(OSDs) that provide Ceph with data durability and integrity using replication,
erasure-coding, and scrubbing~\cite{weil_rados_2007}. Ceph already provides
some degree of programmability; the OSDs support domain-specific object
interfaces that are implemented by composing existing low-level storage
interfaces that execute atomically. These interfaces are written in C++ and are
statically loaded into the system.

@@

\begin{table}[ht]
\centering
  \begin{tabular}{l|l|l}
    Category & Example & \# \\ \hline
    Logging  & Geographically distribute replicas & 11 \\ \hdashline
    Metadata & Snapshots in the block device OR  & \multirow{2}{*}{74} \\
    Management & Scan extents for file system repair & \\ \hdashline
    Locking  & Grants clients exclusive access & 6 \\ \hdashline
    Other & Garbage collection, reference counting  & 4\\
\end{tabular}
\caption{A variety of RADOS object storage classes exist to expose interfaces
    to applications. \# is the number of methods that use these categories.
}
\label{table:objclasses}
\end{table}

@@

@@

\begin{table*}
\centering
\begin{tabular}{  l | l | l | l }
\multicolumn{4}{c}{\Large \textbf{Common Internal Abstractions}} \\
\multicolumn{4}{c}{} \\
\textbf{Abstraction}                    &
\textbf{Example in Production Systems}  &
\textbf{Example in Ceph}                &
\textbf{Provided Primitives}            \\ \hline
{\it Service Metadata}
  & Zookeeper/Chubby coordination~\cite{hunt_zookeeper_2010,website:chubby}
  & cluster state management~\cite{website:ceph-mon}
  & consensus \& consistency
  \\
{\it Shared Resource}
  & MPI collective IO, burst
  & POSIX metadata protocols
  & serialization \& batching
  \\
{\it Object \& Data}
  & Swift in situ storage/compute~\cite{website:zerocloud}
  & object interface classes~\cite{website:cls-lua}
  & transaction \& atomicity
  \\
Durability
  & S3/Swift interfaces (RESTful API)
  & object store library~\cite{weil_rados_2007}
  & interface to persistence
  \\
Load Balancing
  & VMWare's VM migration~\cite{vmware-drs,gulati:hotcloud2011-cloud-resource-management} 
  & migrate POSIX metadata~\cite{weil:sc2004-dyn-metadata}
  & migration \& sampling
  \\
\end{tabular}
\caption{The same ``internal abstractions" are common across large-scale
systems because they provide primitives that solve general distributed systems
problems.  Here we list examples of what these internal abstractions are used
for in ``production systems" and in Ceph.  Malacology provides these internal
abstractions as interfaces that higher level applications can use; {\it
italicized abstractions} are interfaces contributed in this paper.}
\label{table:examples}
\end{table*}

@@

@@

{\bf File type interface.} The metadata service manages a POSIX file system
hierarchy where files and directories are represented as inode data structures
and expose a POSIX file interface. Application that manage large amounts of
metadata (e.g. users or database snapshots) often require a naming service. We
allow new inode types to be defined such that applications can create
domain-specific interfaces to inodes that may modify locking and capability
policies. We will show this is used in Section~\ref{sec:seq} when we discuss a
distributed shared-log built on Malacology.

{\bf Load balancing interface.} Mantle~\cite{sevilla:sc15-mantle}, a
programmable metadata load balancer, has been implemented on Malacology so that
its infrastructure can enjoy all the properties of existing internal
abstractions.  The load balancing logic is intact but the infrastructure has
been improved to safely load, manage, and persist its load balancing policies.
Composing Mantle using the service metadata and durability interfaces, Mantle
can safely version balancer policies, save balancer policies in the backend
object store and centralize warnings/errors. When combined with file types,
load balancing policies can express a variety of rules such as multi-tentant
and multi-application balancing.

@@

\label{services}
\begin{table}
\centering
\begin{tabular}{  l | l | l }
\textbf{Service}              &
\textbf{Used to...}           &
\textbf{Malacology Interface} \\ \hline
Mantle & load balance metadata     & Load Balancing  \\
       & agree on current balancer & Service Metadata\\
       & control client write lock & Shared Resource \\
       & persist policy logic      & Durability      \\ \hline
ZLog   & implement CORFU           & Object \& Data  \\ 
       & install ZLog interfaces   & Service Metadata\\
       & persist log entries       & Durability      \\ \hline
ZLog   & load balance sequencers   & Load Balancing  \\
Seqr.  & manage access to log tail & Shared Resource \\ 
\end{tabular}
\caption{Mantle and ZLog are built using Malacology interfaces so they inherit the
robustness of the subsystems of the underlying stack.}
\label{table:implementation}
\end{table}

@@

\section{DON'T READ THIS!!! Evaluation}
\label{sec:evaluation}

@@

%\subsubsection{Experiment: Object Class
%Programmability}\label{experiment-object-class-programmability}
%
%Add quote from CORFU paper that talks about what they had to hack
%
%\subsubsection{Experiment: Multi-Client
%Burstiness}\label{experiment-multi-client-burstiness}
%
%\begin{figure}[htbp] \centering
%\includegraphics{figures/caps-delay-thruput.png} \caption{Forcing the client
%to drop their capabilities later (delay) improves throughput} \end{figure}
